---
title: "p8105_hw3_ys3298"
author: "Yimeng SHANG"
date: "10/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
```

```{r}
library(p8105.datasets)
library(tidyverse)
library(patchwork)
data("instacart")
data("brfss_smart2010")
accel = read_csv("./accel_data.csv") 
```

```{r include=FALSE}
aisle_num =
  instacart %>%
  group_by(aisle) %>%
  summarise(number = n()) %>%
  count() %>%
  as.integer()
  
aisle_max = 
  instacart %>%
  group_by(aisle) %>%
  summarise(number = n()) %>%
  filter(number == max(number)) %>%
  select(-number) %>%
  as.character()
```
# Problem1
## Short description
**The size and structure of the data:** There are `r nrow(instacart)` observations and `r ncol(instacart)` variables in the data. It's a (`r dim(instacart)`) dataframe. Most of the variable types are integers and some are characters. 

**Some key variables:** Product name; aisle; department; order number and so on. 

**illstrative examples of observations:** For the 1st order, there're 8 different kind of products. Eval set is train. In details, the first product is `r instacart[1,11]`, aisle id is `r instacart[1,12]`, department id is `r instacart[1,13]`, aisle is `r instacart[1,14]`, department is `r instacart[1,15]`. For the second observations, aisle id is `r instacart[2,12]`, department id is `r instacart[2,13]`, aisle is `r instacart[2,14]`, department is `r instacart[2,15]` .

**How many aisles are there, and which aisles are the most items ordered from?** There are `r aisle_num` aisles. The most items ordered from `r aisle_max`.

## Make a plot
```{r echo=FALSE}

  instacart %>%
  group_by(aisle) %>%
  summarise(number = n()) %>%
  filter(number >= 10000) %>%
  ggplot(aes(x = reorder(aisle, number), y = number, fill = aisle)) +
  geom_col() + coord_flip() +
  viridis::scale_fill_viridis(option = "viridis" , discrete = TRUE) +
  labs(
    title = "Aisle Plot",
    x = "aisles",
    y = "number"
  )
```

## Make a table1 
```{r echo=FALSE}
baking = 
  instacart %>%
  group_by(product_name) %>%
  filter(aisle == "baking ingredients") %>%
  summarise(order_times = n()) %>%
  arrange(desc(order_times)) %>%
  head(n = 3) %>%
  mutate(group = "baking ingredients") %>%
  select(group, everything())

dog = 
  instacart %>%
  group_by(product_name) %>%
  filter(aisle == "dog food care") %>%
  summarise(order_times = n()) %>%
  arrange(desc(order_times)) %>%
  head(n = 3) %>%
  mutate(group = "dog food care") %>%
  select(group, everything())

vege =
  instacart %>%
  group_by(product_name) %>%
  filter(aisle == "packaged vegetables fruits") %>%
  summarise(order_times = n()) %>%
  arrange(desc(order_times)) %>%
  head(n = 3) %>%
  mutate(group = "packaged vegetables fruits") %>%
  select(group, everything())

merge(baking, dog, all = TRUE) %>%
  merge(vege, all = TRUE) %>%
  knitr::kable()
```


## Make a table2
```{r echo=FALSE}
instacart %>%
  select(product_name, order_hour_of_day, order_dow) %>%
  filter(product_name == "Pink Lady Apples" | product_name == "Coffee Ice Cream") %>%
  group_by(product_name, order_dow) %>%
  summarise(
    mean_hour = mean(order_hour_of_day)
  ) %>%
  mutate(
    mean_hour = as.integer(mean_hour),
    order_dow = recode(order_dow, "0" = "Sun", "1" = "Mon", "2" = "Tue", "3" = "Wed", "4" = "Thur", "5" = "Fri", "6" = "Sat")
  ) %>%
  pivot_wider(
    names_from = "order_dow",
    values_from = "mean_hour"
  ) %>%  
  knitr::kable()
```

# Problem2
format the data to use appropriate variable names;

```{r include=FALSE}
##cleaning
  brfss_OH =
  brfss_smart2010 %>%
  janitor::clean_names() %>%
  filter(topic == "Overall Health") %>%
  mutate(
    response = factor(response, levels = c("Poor", "Fair", "Good", "Very good", "Excellent"), ordered = TRUE )
  )
```

```{r include=FALSE}
location2002 =
  brfss_OH %>%
  filter(year ==  "2002") %>%
  group_by(locationabbr) %>%
  distinct(locationdesc) %>%
  summarise(total_location = n()) %>%
  filter(total_location >= 7) %>%
  arrange(total_location)
location2002$locationabbr

location2010 =
  brfss_OH %>%
  filter(year ==  "2010") %>%
  group_by(locationabbr) %>%
  distinct(locationdesc) %>%
  summarise(total_location = n()) %>%
  filter(total_location >= 7) %>%
  arrange(total_location)
```
In 2002, `r location2002$locationabbr` were observed at 7 or more locations.

In 2010, `r location2010$locationabbr` were observed at 7 or more locations.

```{r echo=FALSE}
## WHY NA.
Averagedf =
  brfss_OH %>%
  filter(response == "Excellent") %>%
  group_by(locationabbr, year) %>%
  summarise(average = mean(data_value)) %>%
  select(year, locationabbr, average)

  brfss_OH %>%
  filter(response == "Excellent") %>%
  group_by(locationabbr, year) %>%
  summarise(average = mean(data_value)) %>%
  select(year, locationabbr, average)%>%
  ggplot(aes(x = year, y = average, color = locationabbr)) +
  geom_line() + 
    labs(
      title = "Spaghetti Plot",
      x = "Year",
      y = "Average data value"
    ) +
  viridis::scale_color_viridis(
    name ="States",
    discrete = TRUE)
```

```{r echo=FALSE}
a =  brfss_OH %>%
  filter(locationabbr == "NY" & year == "2006" ) %>%
  select(locationdesc, response, data_value) %>%
  group_by(locationdesc) %>%
  ggplot(aes(x = response, y = data_value, color = locationdesc)) +
  geom_point(size = 4, alpha = .5) + geom_line() +
  labs(
      title = "Response Plot for 2016",
      x = "Year",
      y = "Average data value"
    ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)

b =  brfss_OH %>%
  filter(locationabbr == "NY" & year == "2010" ) %>%
  select(locationdesc, response, data_value) %>%
  group_by(locationdesc) %>%
  ggplot(aes(x = response, y = data_value, color = locationdesc)) +
  geom_point(size = 4, alpha = .5) + geom_line() +
  labs(
      title = "Response Plot for 2016",
      x = "Year",
      y = "Average data value"
    ) +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE)

a/b

## how to geom_line????
```

# Problem3
```{r include=FALSE}
accel = 
  accel %>%
  janitor::clean_names()

weekend =
  accel %>%
  filter(day == "Sunday" | day == "Saturday") %>%
  mutate( weekday_end = "weekend") %>%
  select(week, day_id, day, weekday_end, everything())

weekday = 
  accel %>%
  filter(day == "Monday" | day == "Tuesday" | day == "Wednesday" | day == "Thursday" | day == "Friday") %>%
  mutate( weekday_end = "weekday") %>%
  select(week, day_id, day, weekday_end, everything())

accel =
  merge(weekend, weekday, all = TRUE) %>%
  arrange(week, day_id) %>%
    mutate_if(is.double, as.integer)
```
For this resulting dataset, there are `r nrow(accel)` observations and `r ncol(accel)` variables. It also includes the week, day ID, weekend or weekday information.

```{r include=FALSE}
accel$sum = rowSums(accel[, 5:144])
accel = select(accel, week, day_id, day, weekday_end, sum, everything())
table = select(accel, week, day_id, day, weekday_end, sum) 
knitr::kable(table)
# cannot knit
#Are any trends apparent?
```
 
```{r}
table %>%
  #group_by(week) %>%
  select(week, day, sum) %>%
  ggplot(aes(x = week, y = sum, color = day)) +
  geom_point(size = 4, alpha = .5) + geom_line() +
  labs(
      title = "24-hour activity time courses",
      x = "Week",
      y = "Sum of a day"
    ) +
  viridis::scale_color_viridis(
    name ="Location",
    discrete = TRUE)
  
```

Accelerometer data allows the inspection activity over the course of the day. Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week. Describe in words any patterns or conclusions you can make based on this graph

